---
title: "Database builder"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Load packages
```{r setup}
## Load Necessary packages
sapply(c("rentrez", "bold", "taxize","taxizedb", "usethis", "tidyverse", "spider", "insect", "ape", "DECIPHER", "ggpubr", "RColorBrewer", "plotly", "ggforce", "seqinr", "shortread", "patchwork", "viridis","ggridges","UpSetR"), require, character.only = TRUE)


# install.packages("devtools")
devtools::install_github("alexpiper/taxreturn")
library(taxreturn)

## Get an NCBI api key for taxize queries
# use_entrez()

# After generating your key set it as ENTREZ_KEY in .Renviron.
# ENTREZ_KEY='1c0a0c4afa28448650a1450662a22c68f208'
#usethis::edit_r_environ()
```

## Download data for all insecta

```{r retrieve sequences, eval=FALSE, include=FALSE}
## Fetch sequences from GenBank 
genbank <- fetchSeqs("Insecta", database="genbank", out.dir="reference/insecta/genbank", downstream="Order", quiet=FALSE, marker="COI OR COI OR COX1 OR COXI", output = "gb-binom",compress=FALSE, cores=1)

## Fetch sequences from BOLD
bold <- fetchSeqs("Insecta", database="bold", out.dir="reference/insecta/bold", downstream="Order",quiet=FALSE, marker="COI-5P", output = "gb-binom",compress=FALSE, cores=3)

## Fetch mitochondrial genomes from genbank
fetchSeqs("Insecta", database="genbank", out.dir="reference/insecta/genbank", quiet=FALSE, marker="mitochondria", output = "gb-binom", compress=TRUE, cores=2)

```

## Download data for Arachnida

```{r retrieve arachnid seqs, eval=FALSE, include=FALSE}
## Fetch sequences from GenBank 
fetchSeqs("Arachnida", database="genbank", out.dir="genbank", downstream="Order", quiet=FALSE, output = "gb-binom", compress=TRUE, cores=1)

## Fetch sequences from BOLD
fetchSeqs("Arachnida", database="bold", out.dir="reference/arachnida/bold", downstream=TRUE, quiet=FALSE, downto="Order", marker="COI-5P", output = "gb-binom",compress=FALSE, cores=1)

```

# Get some outgroups using random subsampling of search
```{r outgroups}
outgroup_classes <- dplyr::bind_rows(taxize::downstream("Arthropoda", db="itis", downto="Class")) %>%
  filter(!taxonname=="Insecta") %>%
  pull(taxonname)
outgroup_phyla <- dplyr::bind_rows(taxize::upstream("Insecta", db="itis", upto="Phylum")) %>%
  pull(taxonname)
outgroup_kingdoms <- c("Bacteria", "Fungi")

outgroups <- c(outgroup_classes, outgroup_phyla, outgroup_kingdoms)

fetchSeqs(outgroups, database="genbank", out.dir="reference/outgroups", quiet=FALSE, output = "gb-binom", subsample = 100, compress=TRUE, cores=1)
```

## Merge and clean sequences

# Warning - need to run again from start to deal with duplicated accessions now fixed

```{r merge and clean}
seqs <- c(readDNAStringSet(list.files("reference/insecta/genbank", pattern = ".fa", full.names = TRUE)),
          readDNAStringSet(list.files("reference/insecta/bold", pattern = ".fa", full.names = TRUE)),
          readDNAStringSet(list.files("reference/arachnida/genbank", pattern = ".fa", full.names = TRUE)),
          readDNAStringSet(list.files("reference/arachnida/bold", pattern = ".fa", full.names = TRUE)),
          readDNAStringSet(list.files("reference/outgroups", pattern = ".fa", full.names = TRUE))
          )

# write out merged seqs
writeXStringSet(seqs, filepath = "reference/mergedseqs.fa.gz", compress=TRUE)

# Dereplicate duplicated accessions
uniqSeqs <- seqs[!duplicated(str_remove(names(seqs), "\\|.*$")),]
writeXStringSet(uniqSeqs, "reference/uniqSeqs.fa.gz", width=5000, compress=TRUE)

## Map to Open tree of life taxonomy, resolving synonyms and filtering bad flags:
# "incertae_sedis,|incertae_sedis$|major_rank_conflict|unplaced|environmental|inconsistent|extinct|hidden|hybrid|not_otu|viral|barren"
download_ott_taxonomy()
resolved <- map_to_ott(uniqSeqs, dir="ott3.2", resolve_synonyms=TRUE, filter_bads = TRUE, remove_na = TRUE, quiet=FALSE)
insect::writeFASTA(resolved, file="reference/resolved.fa.gz", compress=TRUE)

# Filter any further sequences with problem names
names <- names(readFASTA("reference/codon_filt.fa.gz")) %>%
  str_split_fixed(";", n=2) %>%
  as_tibble() %>%
  set_colnames(c("acc", "name"))
rem <- names %>%
  dplyr::filter(str_detect(name,                        "sp\\.|spp\\.|aff\\.|nr\\.|bv\\.|cf\\.|nom\\.|nud\\.|environment|undescribed|unverified|unclassified|uncultured|unidentif|[0-9]|[:punct:]")) %>%
  pull(acc)
resolved  <- subset.DNAbin(resolved, subset = !str_replace(names(resolved), "(?:.(?!;))+$", "") %in% rem)

#build PHMM from midori longest - sequences need to be same length
midori <-  Biostrings::readDNAStringSet("reference/MIDORI_LONGEST_20180221_COI.fasta")
insecta_midori <- as.DNAbin(midori[str_detect(names(midori),pattern=";Insecta;"),])
folmer <- insect::virtualPCR(insecta_midori, up = "TITCIACIAAYCAYAARGAYATTGG", down= "TAIACYTCIGGRTGICCRAARAAYCA",cores=2, rcdown = TRUE, trimprimers = FALSE)

#Filtered was then aligned in MAFFT and manually curated in geneious prime
folmer_curated <-  ape::read.dna("reference/folmer_insecta_fullength_aligned_curated.fa",format="fasta")
model <- aphid::derivePHMM(folmer_curated)

filtered <- clean_seqs(resolved, model, minscore = 400, cores=2, shave=TRUE, maxNs = 0)
insect::writeFASTA(filtered, file="reference/filtered.fa.gz",compress=TRUE)

# Filter for stop codons
codon_filt <- taxreturn::codon_filter(purged)

# flag clusters with mixed taxonomy at different cluster thresholds and taxonomic ranks
seqs <- insect::subset.DNAbin(codon_filt, subset = !duplicated(names(codon_filt)))

thresholds <- rev(seq(0.9, 1, 0.01))
threshlist <- vector("list", length=length(thresholds))

for (i in 1:length(thresholds)){
  threshlist[[i]] <- taxreturn::get_mixed_clusters(
    x = seqs, source="ott3.2", db=NULL,
    rank = c("species","genus","family"),
    threshold = thresholds[i],
    confidence=0.6, quiet = FALSE) 
}
names(threshlist) <- thresholds 
mixed_clusters <- dplyr::bind_rows(threshlist)

mixed_clusters <- vroom::vroom("reference/mixedclusters.csv")

gg.mixed <- mixed_clusters %>%
  group_by(threshold, rank) %>%
  summarise(n=n()) %>%
  ggplot(aes(x=as.factor(threshold), y=n, fill=rank, group=rank)) + 
  geom_bar(stat="identity", position="dodge", colour="black") +
  xlab("Clustering threshold") +
  ylab("Problem Sequences") +
  ggtitle("Mixed clusters") +
  theme_classic() +
  scale_fill_brewer(name = "Taxonomic Rank", palette="Greens") 

# Purge all with mixed genus at 97, with confidence > 0.6  - >0.8 is probably better - Could i justify this with a probability of lowest common anestor?
rem <- mixed_clusters %>% 
  filter(rank=="genus", threshold ==0.97, confidence > 0.8) %>%
  pull(Acc)


purged  <- subset.DNAbin(codon_filt, subset = !str_replace(names(codon_filt), "(?:.(?!;))+$", "") %in% rem)
insect::writeFASTA(purged, file="reference/purged.fa.gz", compress=TRUE)

# Prune large group sizes down to 5
pruned <- prune_groups(purged, maxGroupSize = 5, discardby="length", dedup=TRUE, quiet = FALSE)

# Reformat to complete taxonomic heirarchy
pruned <- readFASTA("reference/pruned.fa.gz")
db <- get_ott_taxonomy(dir="ott3.2")

# Remove accesions that are identical across both 
pruned <- readDNAStringSet("reference/pruned.fa.gz")
pruned <- pruned[!duplicated(str_remove(names(pruned), "\\|.*$")),] 
db <- get_ott_taxonomy(dir="ott3.2")


## STILL HAVING ISSUES WITH THIS!
pruned <- reformat_hierarchy(pruned, db=db, ranks = c("kingdom","phylum", "class", "order", "family", "genus", "species"), quiet=FALSE)
insect::writeFASTA(pruned, file="reference/pruned.fa.gz", compress=TRUE)
```


## Summarise sequences lost at each stage
```{r sequence tracker}
# Create read origins table
origin <- bind_rows(
  #Genbank Insecta
  fasta.index(list.files("reference/insecta/genbank", pattern = ".fa", full.names = TRUE)) %>%
    mutate(seqid = desc %>%
    str_replace(pattern="(\\|)(.*?)(?=$)", replacement="")) %>%
  mutate(origin = "genbank_insecta") %>%
  select(origin, seqid),
  #BOLD Insecta
  fasta.index(list.files("reference/insecta/bold", pattern = ".fa", full.names = TRUE)) %>%
    mutate(seqid = desc %>%
    str_replace(pattern="(\\|)(.*?)(?=$)", replacement="")) %>%
  mutate(origin = "bold_insecta") %>%
  select(origin, seqid),
  #Genbank Arachnida
  fasta.index(list.files("reference/arachnida/genbank", pattern = ".fa", full.names = TRUE)) %>%
    mutate(seqid = desc %>%
    str_replace(pattern="(\\|)(.*?)(?=$)", replacement="")) %>%
  mutate(origin = "genbank_arachnida") %>%
  select(origin, seqid),
  #BOLD Arachnida
  fasta.index(list.files("reference/arachnida/bold", pattern = ".fa", full.names = TRUE)) %>%
    mutate(seqid = desc %>%
    str_replace(pattern="(\\|)(.*?)(?=$)", replacement="")) %>%
  mutate(origin = "bold_arachnida") %>%
  select(origin, seqid),
  #Outgroups
  fasta.index(list.files("reference/outgroups", pattern = ".fa", full.names = TRUE)) %>%
    mutate(seqid = desc %>%
    str_replace(pattern="(\\|)(.*?)(?=$)", replacement="")) %>%
  mutate(origin = "outgroups") %>%
  select(origin, seqid),
) %>%
# Remove sequences imported from BOLD to genbank
  mutate(Duplicated = case_when(
    duplicated(seqid) & str_detect(origin, "bold") ~ TRUE,
    TRUE ~ FALSE
    )) %>%
  filter(!Duplicated) %>%   
  select(-Duplicated)


## summarise number of sequences at each stage and their origins
tracker <- bind_rows(
                  taxreturn::summarise_fasta("reference/mergedseqs.fa.gz", label="merged", origin=origin),
                  taxreturn::summarise_fasta("reference/uniqSeqs.fa.gz", label="unique", origin=origin),
                  taxreturn::summarise_fasta("reference/filtered.fa.gz", label="phmm_filtered", origin=origin),
                  taxreturn::summarise_fasta("reference/resolved.fa.gz", label="resolved", origin=origin),
                  taxreturn::summarise_fasta("reference/purged.fa.gz", label="purged", origin=origin),
                  taxreturn::summarise_fasta("reference/codon_filt.fa.gz", label="codon_filt", origin=origin),
                  taxreturn::summarise_fasta("reference/pruned.fa.gz", label="pruned", origin=origin)
) %>%
  mutate(label = factor(label, levels=c("merged", "unique", "phmm_filtered", "resolved", "purged", "codon_filt", "pruned" ))) %>%
  pivot_longer(cols=starts_with("n"),
               names_to = "Type",
               values_to = "value"
               )

gg.cleaning <- ggplot(tracker, aes(x=label, y=value, group=origin, fill=origin)) +
  geom_bar(stat="identity") +
  facet_wrap(~Type, nrow=2, ncol=1, scales = "free") +
  scale_fill_brewer(palette="Spectral") +
  xlab("Filter stage") +
  ylab("# Sequences") +
  theme_bw() + 
  scale_y_continuous(labels = scales::comma)
```

# Reformat and Merge in inhouse sequences

```{r reformat seqs}
# read in  inhouse sequences
seqs <- readDNAStringSet(list.files("reference/inhouse/", pattern = ".fa", full.names = TRUE))

#USe NCBI to get lineage
db <- get_ncbi_lineage()

#
names(seqs) <- names(seqs) %>%
   str_split_fixed(pattern="_", n=2) %>%
  as_tibble() %>%
  mutate(tax_name = str_replace(V2, "_", " ")) %>%
  mutate(tax_name = case_when(
    tax_name %in% (anti_join(., db, by="tax_name") %>% pull(tax_name)) ~ tax_name %>% str_replace(pattern="(\\ )(.*?)(?=$)", replacement=""),
    !tax_name %in% (anti_join(., db, by="tax_name") %>% pull(tax_name)) ~ tax_name
  )) %>%
  left_join(db, by="tax_name") %>% 
  filter(!duplicated(V1)) %>%
  mutate(genus = case_when(
    is.na(genus) ~ tax_name,
    !is.na(genus) & !str_detect(tax_name, pattern="\\ ") & !tax_name == genus ~ tax_name,
    TRUE ~ genus
  )) %>%
  unite(col=tax, c("kingdom","phylum", "class", "order", "family", "genus","V2"), sep=";" ) %>%
  unite(col=acc, c("V1", "tax_id"), sep="|") %>%
  unite(col = names, c("acc", "tax"), sep=";") %>%
  pull(names)

#filt using phmm
folmer_curated <-  ape::read.dna("reference/folmer_insecta_fullength_aligned_curated.fa", format="fasta")
model <- aphid::derivePHMM(folmer_curated)

filtered <- clean_seqs(seqs, model, minscore = 400, cores=2, shave=TRUE, maxNs = 0)

# Filter for stop codons
codon_filt <- taxreturn::codon_filter(filtered %>% as.character %>% lapply(.,paste0,collapse="") %>% unlist %>% DNAStringSet)

#merge
mergedseqs <- c((insect::readFASTA(file="reference/pruned.fa.gz")), as.DNAbin(codon_filt))

insect::writeFASTA(mergedseqs, file="reference/merged_final.fa.gz")
```

# Create reference set for fwhf2-fwhR2n region of COI

## Trim to primer regions

```{r Create taxonomic classifier database}
mergedSeqs <- insect::readFASTA("reference/merged_final.fa.gz")
#Trim to primer region using virtualPCR from insect package - using BF1 and BR1 primers
amplicon <- virtualPCR(mergedseqs, up = "ACWGGWTGRACWGTNTAYCC",down= "ARYATDGTRATDGCHCCDGC", cores=2, rcdown = TRUE, trimprimers = TRUE)
writeFASTA(amplicon,"reference/merged_final_bftrimmed.fa.gz")
```

# evaluate coverage


# Evaluate LCA stats

```{r evaluate LCA}

seqs <- readFASTA("reference/pruned.fa.gz")

out <- taxreturn::lca_probs(x, sim=0.95)

pestlist <- read_csv("primer_evaluation/pestlist.csv") %>%
  pull(Family) %>%
  unique()
seqs <- readFASTA("reference/merged_final.fa.gz")

out <- taxreturn::lca_probs(seqs, sim=seq(0.8,1,0.01))

out <- out %>% filter(sim > 50)
gg.lca  <-  ggplot(out, aes(x=sim, y=prob, group=rank, fill=rank, colour=rank)) + 
  geom_line(size=1) +
  geom_vline(xintercept=97, linetype="dotted") +
  scale_x_reverse() +
  scale_color_brewer(palette = "Spectral")  +
  theme_classic()  +
  theme(legend.position = "bottom") +
  xlab("% Sequence Identity") +
  ylab("P(LCR)") + 
  ggtitle("Probability of sequence sharing LCR with reference")


```

# Output trained classifiers


```{r train and output}
amplicon <- insect::readFASTA("reference/old/merged_final_bftrimmed.fa.gz")

#Format for RDP classifier
rdp_genus <- reformat_dada2_gen(amplicon, quiet=FALSE)
insect::writeFASTA(rdp_genus, "reference/rdp_genus.fa.gz", compress=TRUE)

rdp_species <- reformat_dada2_spp(amplicon, quiet=FALSE)
insect::writeFASTA(rdp_species, "reference/rdp_species.fa.gz", compress=TRUE)

# Train IDTAXA
trainingSet <- taxreturn::train_idtaxa(amplicon)
#Write out training set
saveRDS(trainingSet, file="reference/idtaxa.rds")
#Summarise database tree
summary <- taxreturn::tax2tree(amplicon, output="treedf")
write_tsv(summary, "reference/database_summary.txt")

```

# Sessioninfo
```{r sessioninfo}
sessionInfo()
```
