---
title: "sample_switch"
author: "Alexander Piper"
date: "20/05/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load packages
```{r setup}
#Set required packages
.cran_packages <- c("tidyverse",
                    "tidymodels",
                    "patchwork", 
                    "vegan", 
                    "seqinr",
                    "ape", 
                    "RColorBrewer",
                    "ggtree", 
                    "castor", 
                    "picante",
                    "skimr",
                    "devtools")
.bioc_packages <- c("dada2",
                    "phyloseq", 
                    "DECIPHER",
                    "Biostrings",
                    "ShortRead", 
                    "philr",
                    "ALDEx2")

.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
   install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)) {
  if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
  BiocManager::install(.bioc_packages[!.inst], ask = F)
}

sapply(c(.cran_packages,.bioc_packages), require, character.only = TRUE)

# Github packages
devtools::install_github("alexpiper/taxreturn")
library(taxreturn)
devtools::install_github("alexpiper/seqateurs")
library(seqateurs)
devtools::install_github("mikemc/speedyseq")
library(speedyseq)
devtools::install_github('ggloor/CoDaSeq/CoDaSeq')
library(CoDaSeq)
devtools::install_github("mikemc/metacal")
library(metacal)

#Source internal functions
source('R/helper_functions.R')
options(stringsAsFactors = FALSE)
```

# Drosophila 

## Make Phyloseq object

```{r create PS, eval = FALSE}
## Just subset to the last run
seqtab <- readRDS("output/rds/seqtab_final.rds")
#seqtab <- seqtab[str_detect(rownames(seqtab), "HLVKYDMXX"),]
seqtab <- seqtab[,colSums(seqtab) > 0]

# Add FCID to run 1 samples
rownames(seqtab)[!str_detect(rownames(seqtab), "Rep") ] <- paste0("CB3DR_", rownames(seqtab)[!str_detect(rownames(seqtab), "Rep") ])

# Add FCID to run2 samples
rownames(seqtab)[str_detect(rownames(seqtab), "Rep") & 
                   !str_detect(rownames(seqtab), "HLVKYDMXX") & 
                   !str_detect(rownames(seqtab), "DL|CL")] <- paste0("CK3HD_", rownames(seqtab)[str_detect(rownames(seqtab), "Rep") & 
                   !str_detect(rownames(seqtab), "HLVKYDMXX") & 
                   !str_detect(rownames(seqtab), "DL|CL")])

	
# Add FCID to run 3 samples
rownames(seqtab)[str_detect(rownames(seqtab), "Rep") & 
                   !str_detect(rownames(seqtab), "HLVKYDMXX") & 
                   str_detect(rownames(seqtab), "DL|CL")] <- paste0("CJKFJ_", rownames(seqtab)[str_detect(rownames(seqtab), "Rep") & 
                   !str_detect(rownames(seqtab), "HLVKYDMXX") & 
                   str_detect(rownames(seqtab), "DL|CL")])

#Reformat sample IDs
rownames(seqtab)  <- rownames(seqtab) %>%
  str_remove("\\..*$") %>%
  str_replace("\\_S[0-9].*\\_...", replacement="_") %>%
  str_replace("_$", "_1") %>%
  str_replace("1in10", "1:10")

tax <- readRDS("output/rds/tax_IdTaxaExact.rds") 
seqs <- DNAStringSet(colnames(seqtab))
names(seqs) <- seqs
phy <- readRDS("output/rds/phytree.rds")$tree

##### Rename problematic samples
rownames(seqtab)  <- rownames(seqtab) %>%
 str_replace_all("D250M1-", "D250M4REP-") %>% # Works
 str_replace_all("D250M4-", "D250M2REP-") %>% # Works
 str_replace_all("D250M5-", "D250M3REP-") %>% #FAILED library
 str_replace_all("D250M3-", "D250M1REP-") %>% #FP suzukii - low reads
 str_replace_all("D250M2-", "D250M5REP-") %>% #Works
 str_replace_all("D500M1-", "D500M4REP-") %>% #Works 
 str_replace_all("D500M4-", "D500M1REP-") %>% #FP Suzukii
 str_replace_all("D500M5-", "D500M2REP-") %>% #Works
 str_replace_all("D500M3-", "D500M3REP-") %>% #Works but low reads for Suz + Biarmipes 
 str_replace_all("D500M2-", "D500M5REP-") %>% #Works
 str_replace_all("D1000M1-", "D1000M3REP-") %>% #Works
 str_replace_all("D1000M4-", "D1000M1REP-") %>% #Works
 str_replace_all("D1000M5-", "D1000M2REP-") %>% #Works
 str_replace_all("D1000M3-", "D1000M5REP-") %>% #Works
 str_replace_all("D1000M2-", "D1000M4REP-") %>% #Works
 str_replace_all("CM10-", "CM9REP-") %>%
 str_replace_all("CM11-", "CM10REP-") %>%
 str_replace_all("CM9-", "CM11REP-") %>%
 str_replace_all("CML2-", "CML6REP-")%>%
 str_replace_all("CML3-", "CML2REP-")%>%
 str_replace_all("CML4-", "CML3REP-")%>%
 str_replace_all("CML5-", "CML4REP-")%>%
 str_replace_all("CML6-", "CML5REP-")%>%
 str_replace_all("CT5-", "CT4REP-")%>%
 str_replace_all("CT4-", "CT5REP-") %>%
str_replace_all("REP", "")

#Load sample information
## ---- samdat ----
samdf <- read.csv("sample_data/Sample_info.csv", header=TRUE) %>% 
  mutate(Sample_ID = case_when(
    FCID=="HLVKYDMXX" ~ paste0(Sample_Name, "_", replicate),
    !FCID=="HLVKYDMXX" ~ paste0(FCID, "_", Sample_Name, "_", replicate)
  ))  %>%
  filter(!(index=="ATCGATCG" & index2=="ATCACACG"), #CT11-ex1 duplicated
         !(index=="TCGCTGTT" & index2=="ACTCCATC") # CT12-ex1 duplicated
         ) %>%
  filter(!FCID=="CK3HD") %>%
  mutate(type = case_when(
    str_detect(Sample_ID, "D[0-9][0-9][0-9]M|D[0-9][0-9][0-9][0-9]M|DM[1-5]-")  ~ "DrosMock",
    str_detect(Sample_ID, "SPD")  ~ "SPD",
    str_detect(Sample_ID, "ACV")  ~ "ACV",
    str_detect(Sample_ID, "DC")  ~ "DC",
    str_detect(Sample_ID, "Sach")  ~ "Sachet",
    str_detect(Sample_ID, "FF")  ~ "FF",
    str_detect(Sample_ID, "NTC")  ~ "NTC",
    str_detect(Sample_ID, "DLarv")  ~ "DrosLarv",
    str_detect(Sample_ID, "POS|SynMock")  ~ "POS",
    str_detect(Sample_ID, "extblank|BLANK")  ~ "Extblank",
    str_detect(Sample_ID, "pcrblank")  ~ "PCRblank",
    str_detect(Sample_ID, "CT")  ~ "CarpTrap",
    str_detect(Sample_ID, "CM")  ~ "CarpMock"
  )) %>%
  mutate(target_subfragment = case_when(
    str_detect(Fprimer, "GGDACWGGWTGAACWGTWTAYCCHCC") & str_detect(Rprimer, "GTRATWGCHCCDGCTARWACWGG") ~ "fwhF2-fwhR2n",
    str_detect(Fprimer, "ACWGGWTGRACWGTNTAYCC") & str_detect(Rprimer, "ARYATDGTRATDGCHCCDGC") ~ "BF1-BR1",
    str_detect(Fprimer, "GGDRCWGGWTGAACWGTWTAYCCNCC") & str_detect(Rprimer, "TATDGTRATDGCHCCNGC") ~ "SauronS878-HexCOIR4",
    str_detect(Fprimer, "GGDACWGGWTGAACWGTWTAYCCHCC") & str_detect(Rprimer, "TATDGTRATDGCHCCNGC") ~ "fwhF2-HexCOIR4",
  )) %>%
  magrittr::set_rownames(.$Sample_ID)


test <- samdf %>% 
  dplyr::select(Sample_Name, type)
# Will probably need to rename the seqtabs and append the flowcell number onto the samples before they are merged

## ---- phyloseq ----
ps <- phyloseq(tax_table(tax), 
               sample_data(samdf),
               otu_table(seqtab, taxa_are_rows = FALSE),
               phy_tree(phy),
               refseq(seqs))

if(nrow(seqtab) > nrow(sample_data(ps))){warning("Warning: All samples not included in phyloseq object, check sample names match the sample metadata")}

rownames(samdf)[which(!rownames(sample_data(ps))  %in% rownames(samdf))]
rownames(sample_data(ps))[which(!rownames(samdf)  %in% rownames(sample_data(ps)))]

# Rename all taxa
taxa_names(ps) <- paste0("SV", seq(ntaxa(ps)),"-",tax_table(ps)[,7])

saveRDS(ps, "output/rds/ps_idtaxaExact.rds") 

#Rename synthetic orders
tax_table(ps)[,2][which(str_detect(tax_table(ps)[,7], "Synthetic"))] <- "Arthropoda"
```


## Taxon filtering

```{R taxon filt}
get_taxa_unique(ps, "Order")

ps # Check the number of taxa prior to removal
ps0 <- ps %>%
  subset_taxa(
    Phylum == "Arthropoda" & 
    Class %in% c("Insecta", "Arachnida", "Collembola")
  )
ps # Confirm that the taxa were removed
get_taxa_unique(ps0, "Phylum")
get_taxa_unique(ps0, "Class")
get_taxa_unique(ps0, "Order")
```

# Merge technical replicates

```{r merge replicates}
# Merge replicates
ps.merged <- ps0 %>%
    merge_samples(group = "Sample_Name", fun="sum")

#This loses the sample metadata - Need to add it agian
sample_data(ps.merged) <- sample_data(ps0) %>%
  filter(!duplicated(Sample_Name)) %>%
  magrittr::set_rownames(.$Sample_Name)

ps1 <- ps.merged
```

# Bias permutations

```{r bias}
ps_bias <- ps1 %>%
  subset_samples(target_subfragment == "fwhF2-fwhR2n") %>%
  #subset_samples(type %in% c("CarpMock",  "CarpTrap")) %>%
  subset_samples(type %in% c("DrosMock")) %>%
  subset_samples(FCID %in% c("HLVKYDMXX", "CB3DR")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE) 

tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Carpophilus_dimidiatus/nr.dimidiatus")] <- "Carpophilus_nr.dimidiatus"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp1")] <- "Brachypeplus_Sp"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp2")] <- "Brachypeplus_Sp"
ps_bias <- speedyseq::tax_glom(ps_bias, taxrank="Species")

exp <- read_csv("sample_data/expected_quant_merged.csv") %>%
  pivot_longer(-X1,
               names_to= "Species",
               values_to= "Abundance") %>%
  mutate(Species = str_replace(Species, pattern=" ",replacement="_")) %>%
  mutate(X1 = str_split_fixed(X1,"-rep",n=2) %>%
           as_tibble()%>% 
           pull(V1)) %>%
  drop_na() %>%
  magrittr::set_colnames(c("Sample","Taxon","Actual"))%>%
  filter(str_detect(Sample, "D100M")) %>% #Use just the 100 abundance compositions
  #filter(str_detect(Sample, "D[0-9][0-9][0-9]M|D[0-9][0-9][0-9][0-9]M")) %>%
  mutate(comm = str_extract(Sample,"M[0-9]")) %>%
  dplyr::select(-Sample)

sam <- speedyseq::psmelt(ps_bias) %>%
  arrange(Abundance) %>%
  mutate(Taxon = Species) %>%
  filter(!str_detect(Genus, "__")) %>%#Remove unclassified
  mutate(comm = case_when(
    type=="DrosMock" ~ str_extract(Sample_Name,"M[0-9]"),
    TRUE ~ as.character(NA) )) %>% 
  dplyr::select(Sample_Name,Taxon, Abundance, target_subfragment, FCID, comm) %>%
  mutate(Sample = Sample_Name %>%
           str_remove_all("BF1-BR1-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-|fwhF2-fwhR2n-") %>%
           str_remove("HLVKYDMXX_") %>%
           str_remove("-ex[0-9]"))%>%
  filter(Taxon %in% exp$Taxon) %>%
  #filter(Sample %in% exp$Sample) %>%
  filter(str_detect(Taxon, "Dros")) %>%
  mutate(commsize = Sample %>% 
           str_remove(pattern="M[0-9]$") %>%
           str_remove(pattern="^D")) %>%
  mutate(commsize = case_when(
    str_detect(Sample_Name, "DM[1-9]-")  ~ as.character(100),
    TRUE ~ commsize
  ))


# Get all permutations
set.seed(666)  
elements <- unique(sam$comm)

#Allow for 2 M3's for run 500 to deal with failed sample

k <- length(elements) 
#for run 500 Allow duplcated m3 c(elements, "M3")
res <- purrr::map(1:1000, function(x) sample(elements, k, replace=FALSE)) %>%
  unique() %>%
  purrr::map(function(x){
    tibble(comm=elements, newcomm=x)})
# below, check you have all the permutations you need (if not, try again)
length(res) == factorial(k)

# Split into known and mixed samples
sam2 <- sam %>% 
  filter(FCID=="HLVKYDMXX") %>%
  group_by(commsize) %>%
  group_split()

test <- sam2[[4]]

permuted <- res %>%
  purrr::map(function(x){ 
    df <- x %>% 
      full_join(test, by="comm") %>%
      dplyr::select(-comm) %>%
      dplyr::rename(comm = newcomm) %>% 
      bind_rows(sam %>% 
       filter(!commsize=="500")) %>% # For run  500
        #filter(FCID=="CB3DR")) %>% #Bind to known
      left_join(exp, by = c("comm","Taxon")) %>%
      mutate(Actual = replace_na(Actual, 0)) %>%
      mutate(Observed0 = (Abundance + 0.5) * (Actual > 0)) %>%
      mutate(Error = Observed0 / Actual) %>%
      distinct() 
    
    #Calculate bias
    bias <- df %>%
    filter(Actual > 0 ) %>%
    group_by(Taxon) %>%
    summarize(Bias = gm_mean(Error), se=gm_sd(Error)) %>%
    mutate(Bias = center_elts(Bias)) %>%
    ungroup
    
    # Sumamrise bias
    out <- list(
      df = df,
      bias = bias,
      se = bias %>%
      summarise(se = sum(se)),
      transtab=x)
    return(out)
    } )

# Rank SE's  - Get the permutation that minimises se
permut_ranks <- permuted %>%
  map("se") %>%
  bind_rows(.id="permutation") %>%
  dplyr::arrange(se)

permut_ranks

top_permut <- permut_ranks %>%
  dplyr::slice(1) %>%
  pull(permutation) %>%
  as.numeric()

# Visualise the error of the top permutation in all pairwise ratios
permuted[[top_permt]]$df %>%
  filter(Actual > 0) %>%
    compute_ratios(group_vars = c("Sample_Name", "FCID")) %>%
    filter(!is.na(Actual)) %>%
    mutate(Pair = paste(Taxon.x, Taxon.y, sep = ":")) %>% 
  ggplot(aes(Pair, Error, colour=Sample_Name)) +
  geom_jitter(alpha=0.7) +
  geom_hline(yintercept = 1) +
  scale_y_log10() +
  facet_wrap(~FCID)+
  theme_bw()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=45, hjust=1))

#Check the translation table ro the best match
permuted[[top_permut]]$transtab 

translation <- permuted[[top_permut]]$transtab %>%
  full_join(test)
# May need to make it robust to low abundance d.suzukii, d.biarmipes, d. subpulchrella?
```


# Heatmap
```{r check spikeins}
gg.hmap <- plot_heatmap(ps_bias, "jaccard", "jsd", taxa.label="Species", na.value=NA, taxa.order="Family")   +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.spacing =unit(0.2, "lines"),
        legend.position = "bottom",
        legend.direction = "horizontal") +
  facet_grid(~target_subfragment,scales="free",space="free", drop=TRUE) 

gg.hmap
```


## Bias whole dataset

```{r Drosophila bias}
ps_bias <- ps1 %>%
  subset_samples(target_subfragment == "fwhF2-fwhR2n") %>%
  subset_samples(type %in% "DrosMock") %>%
  subset_samples(FCID %in% c("HLVKYDMXX", "CB3DR")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE) 

tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Carpophilus_dimidiatus/nr.dimidiatus")] <- "Carpophilus_nr.dimidiatus"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp1")] <- "Brachypeplus_Sp"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp2")] <- "Brachypeplus_Sp"
ps_bias <- speedyseq::tax_glom(ps_bias, taxrank="Species")

exp <- read_csv("sample_data/expected_quant_merged.csv") %>%
  pivot_longer(-X1,
               names_to= "Species",
               values_to= "Abundance") %>%
  mutate(Species = str_replace(Species, pattern=" ",replacement="_")) %>%
  mutate(X1 = str_split_fixed(X1,"-rep",n=2) %>%
           as_tibble()%>% 
           pull(V1)) %>%
  drop_na() %>%
  magrittr::set_colnames(c("Sample","Taxon","Actual"))

sam <- speedyseq::psmelt(ps_bias) %>%
  arrange(Abundance) %>%
  mutate(Taxon = Species) %>%
  filter(!str_detect(Genus, "__")) %>%
  dplyr::select(Sample_Name,Taxon, Abundance, target_subfragment, FCID) %>%
  mutate(Sample = Sample_Name %>%
           str_remove_all("BF1-BR1-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-|fwhF2-fwhR2n-") %>%
           str_remove("HLVKYDMXX_") %>%
           str_remove("-ex[0-9]"))%>%
  filter(Taxon %in% exp$Taxon) %>%
  filter(str_detect(Taxon, "Dros"))

#Join tables 
joint <- sam %>%
  left_join(exp, by = c("Sample","Taxon")) %>%
  mutate(Actual = replace_na(Actual, 0)) %>%
  mutate(Observed0 = (Abundance + 0.5) * (Actual > 0)) %>%
  mutate(Error = Observed0 / Actual) %>%
  distinct() %>%
  group_by(Sample_Name) %>%
  mutate_at(vars(Abundance, Actual), ~ . / sum(.) ) %>% #Convert to proportions
  dplyr::select(Sample_Name, Taxon, Abundance, Actual, Error, FCID) %>%
  ungroup()

 
## Visualise errors in proportions
library(ggbeeswarm)
library(ggthemes)
errorplot <- joint %>%
  filter(Actual > 0) %>%
    compute_ratios(group_vars = c("Sample_Name", "FCID")) %>%
    filter(!is.na(Actual)) %>%
    mutate(Pair = paste(Taxon.x, Taxon.y, sep = ":")) %>% 
  ggplot(aes(Pair, Error, colour=Sample_Name)) +
  geom_jitter(alpha=0.7) +
  geom_hline(yintercept = 1) +
  scale_y_log10() +
  facet_wrap(~FCID)+
  theme_bw()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=45, hjust=1))

errorplot

library(plotly)
ggplotly(errorplot)


#Build error matrix
error_mat <- build_matrix(joint, Sample_Name, Taxon, Error) 
rows_to_keep <- rowSums(error_mat, na.rm=TRUE) > 0
cols_to_keep <- colSums(error_mat, na.rm=TRUE) > 0
error_mat <- error_mat[rows_to_keep,cols_to_keep]

#Estimate bias using metacal
bias <- center(error_mat, enframe = TRUE) %>%
    dplyr::rename(Bhat = Center)

#Estimate uncertainty in bias estimate using bootstrapping
bootreps <- bootrep_center(error_mat) %>%
    dplyr::rename(Bhat = Center)
bootreps.summary <- bootreps %>%
    group_by(Taxon) %>%
    summarize(Gm_mean = gm_mean(Bhat), Gm_se = gm_sd(Bhat))
bias0 <- left_join(bias, bootreps.summary, by = "Taxon")

```

# Carpophilus

# Bias permutations

```{r bias}
ps_bias <- ps1 %>%
  subset_samples(target_subfragment == "fwhF2-fwhR2n") %>%
  subset_samples(type %in% c("CarpMock",  "CarpTrap")) %>%
  subset_samples(FCID %in% c("HLVKYDMXX", "CB3DR")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE) 

tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Carpophilus_dimidiatus/nr.dimidiatus")] <- "Carpophilus_nr.dimidiatus"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp1")] <- "Brachypeplus_Sp"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp2")] <- "Brachypeplus_Sp"
ps_bias <- speedyseq::tax_glom(ps_bias, taxrank="Species")

exp <- read_csv("sample_data/expected_quant_carpophilus.csv") %>%
  pivot_longer(-X1,
               names_to= "Species",
               values_to= "Abundance") %>%
  mutate(Species = str_replace(Species, pattern=" ",replacement="_")) %>%
  mutate(X1 = str_split_fixed(X1,"-rep",n=2) %>%
           as_tibble()%>% 
           pull(V1)) %>%
  drop_na() %>%
  magrittr::set_colnames(c("Sample","Taxon","Actual"))%>%
  filter(str_detect(Sample, "CM[0-9]|CT[0-9]|CML[0-9]")) %>%
  mutate(Sample = Sample %>% str_remove("-ex[0-9]")) %>%
  distinct()

exp %>% 
  group_by(Sample) %>%
  mutate_at(vars(Actual), ~ . / sum(.) ) %>% #Convert to proportions
  ggplot(aes(x=Sample, y=Actual, fill=Taxon)) +
  geom_col(position="stack")

sam <- speedyseq::psmelt(ps_bias) %>%
  arrange(Abundance) %>%
  mutate(Taxon = Species) %>%
  filter(!str_detect(Genus, "__")) %>%#Remove unclassified
  dplyr::select(Sample_Name,Taxon, Abundance, target_subfragment, FCID) %>%
  mutate(Sample = Sample_Name %>%
           str_remove_all("BF1-BR1-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-|fwhF2-fwhR2n-") %>%
           str_remove("HLVKYDMXX_") %>%
           str_remove("-ex[0-9]"))%>%
  filter(Taxon %in% exp$Taxon,
         Sample %in% exp$Sample) %>%
  mutate(chunk = case_when(
    str_detect(Sample, "CM[1-5]$") ~ "CM[1-5]",
    str_detect(Sample, "CM[6-9]$|CM10$|CM11$") ~ "CM[6-11]",
    str_detect(Sample, "CML[1-6]") ~ "CML[1-6]",
    str_detect(Sample, "CT[1-6]$") ~ "CT[1-6]$",
    str_detect(Sample, "CT[7-9]$|CT10$|CT11$|CT12$") ~ "CT[7-12]"
    ))

# Split into known and mixed samples
known <- sam %>% 
  #filter(FCID=="CB3DR")
  filter(chunk %in% c("CM[1-5]","CT[7-12]"))

mixed <- sam %>% 
  filter(FCID=="HLVKYDMXX") %>%
  filter(!chunk %in% c("CM[1-5]","CT[7-12]"))%>%
  group_by(chunk) %>%
  group_split()

#Subset to test set
test <- mixed[[3]]
  
# Get all permutations for each chunk
set.seed(666)  
elements <- unique(test$Sample)
k <- length(elements) 
#for run 500 Allow duplcated m3 c(elements, "M3")
res <- purrr::map(1:100000, function(x) sample(elements, k, replace=FALSE)) %>%
  unique() %>%
  purrr::map(function(x){
    tibble(Sample=elements, new_sample=x)})
# below, check you have all the permutations you need (if not, try again)
length(res) == factorial(k)

permuted <- res %>%
  purrr::map(function(x){ 
    df <- x %>% 
      full_join(test, by="Sample") %>%
      dplyr::select(-Sample) %>%
      dplyr::rename(Sample = new_sample) %>% 
      bind_rows(known) %>% #joint to known samples
      left_join(exp, by = c("Sample","Taxon")) %>%
      mutate(Actual = replace_na(Actual, 0)) %>%
      mutate(Observed0 = (Abundance + 0.5) * (Actual > 0)) %>%
      mutate(Error = Observed0 / Actual) %>%
      distinct() 
    
    #Calculate bias
    bias <- df %>%
    filter(Actual > 0 ) %>%
    group_by(Taxon) %>%
    summarize(Bias = gm_mean(Error), se=gm_sd(Error)) %>%
    mutate(Bias = center_elts(Bias)) %>%
    ungroup
    
    # Sumamrise bias
    out <- list(
      df = df,
      bias = bias,
      se = bias %>%
      summarise(se = sum(se)),
      transtab=x)
    return(out)
    } )

# Rank SE's  - Get the permutation that minimises se
permut_ranks <- permuted %>%
  map("se") %>%
  bind_rows(.id="permutation") %>%
  dplyr::arrange(se)

permut_ranks

## current observed arrangement SE
test %>% 
      bind_rows(known) %>% #joint to known samples
      left_join(exp, by = c("Sample","Taxon")) %>%
      mutate(Actual = replace_na(Actual, 0)) %>%
      mutate(Observed0 = (Abundance + 0.5) * (Actual > 0)) %>%
      mutate(Error = Observed0 / Actual) %>%
      distinct() %>%
    filter(Actual > 0 ) %>%
    group_by(Taxon) %>%
    summarize(Bias = gm_mean(Error), se=gm_sd(Error)) %>%
    mutate(Bias = center_elts(Bias)) %>%
    ungroup%>%
      summarise(se = sum(se))

top_permut <- 22

# Visualise the error of the top permutation in all pairwise ratios
permuted[[top_permut]]$df %>%
  filter(Actual > 0) %>%
    compute_ratios(group_vars = c("Sample_Name", "FCID")) %>%
    filter(!is.na(Actual)) %>%
    mutate(Pair = paste(Taxon.x, Taxon.y, sep = ":")) %>% 
  ggplot(aes(Pair, Error, colour=Sample_Name)) +
  geom_jitter(alpha=0.7) +
  geom_hline(yintercept = 1) +
  scale_y_log10() +
  facet_wrap(~FCID)+
  theme_bw()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=45, hjust=1))


#Check the translation table ro the best match
permuted[[top_permut]]$transtab 

translation <- permuted[[top_permut]]$transtab %>%
  full_join(test)
# May need to make it robust to low abundance d.suzukii, d.biarmipes, d. subpulchrella?
```

#Rename carpophilus
```{R}
rownames(seqtab)  <- rownames(seqtab) %>%
 str_replace_all("CM10-", "CM9REP-") %>%
 str_replace_all("CM11-", "CM10REP-") %>%
 str_replace_all("CM9-", "CM11REP-") %>%
 str_replace_all("CML2-", "CML6REP-")%>%
 str_replace_all("CML3-", "CML2REP-")%>%
 str_replace_all("CML4-", "CML3REP-")%>%
 str_replace_all("CML5-", "CML4REP-")%>%
 str_replace_all("CML6-", "CML5REP-")%>%
 str_replace_all("CT5-", "CT4REP-")%>%
 str_replace_all("CT4-", "CT5REP-")

test2 <- sam %>%
  group_by(Sample_Name) %>%
  mutate_at(vars(Abundance), ~ . / sum(.) ) %>%#Convert to proportions
  mutate(Abundance = round(Abundance, 3))
```
Chunk1- No change
Sample new_sample 
CM1      CM1
CM2      CM2
CM3      CM3
CM4      CM4
CM5      CM5
Chunk2 - 
  Sample new_sample   
1 CM10 -> CM9  
2 CM11 -> CM10    
3 CM6    CM6   
4 CM9 -> CM11      
5 CM7    CM7     
Chunk3- Larval - no clear winner
 Sample new_sample
1 CML2 ->   CML6  #Makes sense
2 CML3 ->   CML2  # Makes sense
3 CML4 ->   CML3  # CMl5 has more hemi?
4 CML5 ->   CML4  # makes sense due to lack of hemi    
5 CML6 ->   CML5  # Makes sense
6 CML1   CML1  

Chunk4- Traps - No clear winner
1 CT1    CT1 - CT1 an outlier
2 CT3    CT3       
3 CT5 ->   CT4       
4 CT4 ->   CT5  
Chunk5 - Traps no change
  Sample new_sample
  <chr>  <chr>     
1 CT8    CT8       
2 CT11   CT11      
3 CT10   CT10      
4 CT12   CT12      
5 CT9    CT9   


# Heatmap
```{r check spikeins}
gg.hmap <- plot_heatmap(ps_bias, "jaccard", "jsd", taxa.label="Species", na.value=NA, taxa.order="Family")   +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.spacing =unit(0.2, "lines"),
        legend.position = "bottom",
        legend.direction = "horizontal") +
  facet_grid(~target_subfragment,scales="free",space="free", drop=TRUE) 

gg.hmap
```


## Bias whole dataset

```{r Drosophila bias}
ps_bias <- ps1 %>%
  subset_samples(target_subfragment == "fwhF2-fwhR2n") %>%
  subset_samples(type %in% "DrosMock") %>%
  #subset_samples(type %in% c("CarpMock",  "CarpTrap")) %>%
  subset_samples(FCID %in% c("HLVKYDMXX", "CB3DR")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE) 

tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Carpophilus_dimidiatus/nr.dimidiatus")] <- "Carpophilus_nr.dimidiatus"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp1")] <- "Brachypeplus_Sp"
tax_table(ps_bias)[,7][which(tax_table(ps_bias)[,7]=="Brachypeplus_Sp2")] <- "Brachypeplus_Sp"
ps_bias <- speedyseq::tax_glom(ps_bias, taxrank="Species")

exp <- read_csv("sample_data/expected_quant_merged.csv") %>%
  pivot_longer(-Sample,
               names_to= "Species",
               values_to= "Abundance") %>%
  mutate(Species = str_replace(Species, pattern=" ",replacement="_")) %>%
  mutate(Sample = str_split_fixed(Sample,"-rep",n=2) %>%
           as_tibble()%>% 
           pull(V1)) %>%
  drop_na() %>%
  magrittr::set_colnames(c("Sample","Taxon","Actual"))

sam <- speedyseq::psmelt(ps_bias) %>%
  arrange(Abundance) %>%
  mutate(Taxon = Species) %>%
  filter(!str_detect(Genus, "__")) %>%
  dplyr::select(Sample_Name,Taxon, Abundance, target_subfragment, FCID) %>%
  mutate(Sample = Sample_Name %>%
           str_remove_all("BF1-BR1-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-|fwhF2-fwhR2n-") %>%
           str_remove("HLVKYDMXX_") %>%
           str_remove("-ex[0-9]"))%>%
  filter(Taxon %in% exp$Taxon) 

#Join tables 
joint <- sam %>%
  left_join(exp, by = c("Sample","Taxon")) %>%
  mutate(Actual = replace_na(Actual, 0)) %>%
  mutate(Observed0 = (Abundance + 0.5) * (Actual > 0)) %>%
  mutate(Error = Observed0 / Actual) %>%
  distinct() %>%
  group_by(Sample_Name) %>%
  mutate_at(vars(Abundance, Actual), ~ . / sum(.) ) %>% #Convert to proportions
  dplyr::select(Sample_Name, Taxon, Abundance, Actual, Error, FCID) %>%
  ungroup()

 
## Visualise errors in proportions
library(ggbeeswarm)
library(ggthemes)
errorplot <- joint %>%
  filter(Actual > 0) %>%
    compute_ratios(group_vars = c("Sample_Name", "FCID")) %>%
    filter(!is.na(Actual)) %>%
    mutate(Pair = paste(Taxon.x, Taxon.y, sep = ":")) %>% 
  ggplot(aes(Pair, Error, colour=Sample_Name)) +
  geom_jitter(alpha=0.7) +
  geom_hline(yintercept = 1) +
  scale_y_log10() +
  facet_wrap(~FCID)+
  theme_bw()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=45, hjust=1))

errorplot

library(plotly)
ggplotly(errorplot)


#Build error matrix
error_mat <- build_matrix(joint, Sample_Name, Taxon, Error) 
rows_to_keep <- rowSums(error_mat, na.rm=TRUE) > 0
cols_to_keep <- colSums(error_mat, na.rm=TRUE) > 0
error_mat <- error_mat[rows_to_keep,cols_to_keep]

#Estimate bias using metacal
bias <- center(error_mat, enframe = TRUE) %>%
    dplyr::rename(Bhat = Center)

#Estimate uncertainty in bias estimate using bootstrapping
bootreps <- bootrep_center(error_mat) %>%
    dplyr::rename(Bhat = Center)
bootreps.summary <- bootreps %>%
    group_by(Taxon) %>%
    summarize(Gm_mean = gm_mean(Bhat), Gm_se = gm_sd(Bhat))
bias0 <- left_join(bias, bootreps.summary, by = "Taxon")

bias0

ggplot(bias0, aes(x=Taxon, y=Gm_mean-1)) + 
  geom_col() +
  geom_pointrange(aes(y = Gm_mean-1, 
            ymin = Gm_mean-1 / Gm_se^2, ymax = Gm_mean-1 * Gm_se^2))
```
