---
title: "Drosophila Metabarcoding"
title: "Statistical analysis"
author: "Alexander Piper"
date: "`r Sys.Date()`"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Introduction

## Analysis structure

Part 1 - Cleanup & Filtering of index switching

Part 1 - Comparison of 4 primers for bias and detection efficiency (figure 1)

Part 2 - Comparison of bias between 3 tagged primers

Part 3 - Comparison

## Load packages

```{r setup}
#Set required packages
.cran_packages <- c("tidyverse",
                    "tidymodels",
                    "patchwork", 
                    "vegan", 
                    "seqinr",
                    "ape", 
                    "RColorBrewer",
                    "castor", 
                    "picante",
                    "phytools",
                    "ggrepel",
                    "devtools")
.bioc_packages <- c("dada2",
                    "phyloseq", 
                    "ggtree",
                    "DECIPHER",
                    "Biostrings",
                    "ShortRead", 
                    "philr",
                    "ALDEx2")

.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
   install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)) {
  if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
  BiocManager::install(.bioc_packages[!.inst], ask = F)
}

sapply(c(.cran_packages,.bioc_packages), require, character.only = TRUE)

# Github packages
devtools::install_github("alexpiper/taxreturn")
devtools::install_github("alexpiper/seqateurs")
devtools::install_github("mikemc/speedyseq")
devtools::install_github('ggloor/CoDaSeq/CoDaSeq')
devtools::install_github("mikemc/metacal")

library(speedyseq)
library(taxreturn)
library(seqateurs)
library(CoDaSeq)
library(metacal)

#Source internal functions
source('R/helper_functions.R')

#Source themes
source('R/themes.R')
```

## Read in phyloseq object

```{r phyloseq}
ps2 <- readRDS("output/rds/ps_filtered.rds")
```

# Run 1 - primer comparison

## Overview of run 1

```{r run 1 overview}
ps_run1 <- ps2 %>%
  subset_samples(fcid %in% c("CB3DR")) 
ps_run1 <- ps_run1 %>%
  subset_samples(!str_detect(sample_names(ps_run1),"blank")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE) 

ps_run1 %>% 
  speedyseq::psmelt() %>%
  group_by(sample_id) %>%
  mutate_at(vars(Abundance), ~ . / sum(.) ) %>% #Convert to proportions
  mutate(plotlabel = case_when(
    Abundance >= 0.01  ~ Species, # Change this to whatever taxrank we want
    Abundance < 0.01 ~ as.character(NA)
    )) %>%
  ggplot(aes(x=sample_name, y=Abundance, fill=plotlabel)) +
  geom_col(position="stack") + 
  facet_grid(pcr_primers~type, scales="free") +
  scale_y_continuous(labels = scales::percent) +
  coord_flip()+
  base_theme+
  theme(legend.position = "bottom") +
  labs(x = "Sample Name",
       y= "Relative abundance",
       fill="Species",
       title = "Run 1 - Primer testing")

# Reads per sample
ps_run1 %>% 
  speedyseq::psmelt() %>%
  group_by(sample_id, sample_name, pcr_primers) %>%
  summarise(Abundance = sum(Abundance)) %>%
  ggplot(aes(x = sample_name, y = Abundance))+
  geom_col() +
  geom_text(aes(label=Abundance),angle=90,hjust=1,vjust=0.5, colour="white")+
  facet_grid(~pcr_primers,scales="free_x", drop=TRUE) + 
  base_theme

ps_run1 %>% 
  speedyseq::psmelt() %>%
  group_by(sample_id, sample_name, pcr_primers) %>%
  summarise(Abundance = sum(Abundance)) %>%
  ungroup() %>%
  summarise(total = sum(Abundance), mean=mean(Abundance), sd = sd(Abundance), upper = range(Abundance)[2], lower=range(Abundance)[1])

# Taxa per primer
ps_run1 %>% 
  speedyseq::psmelt() %>%
  filter(Abundance > 0, !is.na(Abundance)) %>%
  mutate(type = case_when(
    str_detect(sample_name,"D100M") ~ "MOCK",
    TRUE ~ "TRAP"
  ))%>%
  group_by(type, pcr_primers) %>%
  summarise(unique = n_distinct(Species)) 

# Heatmap separate for primers

# Plot phylogeny
exp <- read_csv("sample_data/expected_quant_merged.csv") %>%
  dplyr::rename(sample_name = Sample) %>%
  pivot_longer(-sample_name,
               names_to= "Species",
               values_to= "expected") %>%
  mutate(taxon = str_replace(taxon, pattern=" ",replacement="_"),
         sample_name = str_remove(sample_name, "-exp*.$")) %>%
  dplyr::filter(!is.na(sample_name),
  !str_detect(sample_name, "CM[0-9]|CT[0-9]|CML[0-9]|Syn"),
  expected > 0) %>%
  distinct()

heatmap_dat <-  ps_run1 %>%
  speedyseq::psmelt() %>%
  dplyr::select(sample_id, OTU, sample_name, Abundance, pcr_primers, Species, type) %>%
  filter(!str_detect(Species, "__")) %>%
  mutate(sample_name = sample_name %>% str_remove("CB3DR_"),
         Species = str_replace(Species, "_", " "),
         type = case_when(
           str_detect(sample_name, "D100") ~ "Mock",
           TRUE ~ "Trap"
         )) %>%
  group_by(sample_id) %>%
  mutate_at(vars(Abundance), ~ . / sum(.) ) %>%
  ungroup() %>%  
  mutate(Abundance = na_if(Abundance, 0))  %>%
  left_join(read_csv("sample_data/expected_quant_merged.csv") %>%
  dplyr::rename(sample_name = Sample) %>%
  pivot_longer(-sample_name,
               names_to= "Species",
               values_to= "expected") %>%
    mutate(Species = Species %>% str_replace("_", " ")), by = c("sample_name","Species")) %>%
  mutate(expected = na_if(expected, 0))  %>%
  mutate(outcome = case_when(
    type == "Mock" & Abundance > 0 & is.na(expected) ~ "FP",
    type == "Mock" & Abundance > 0 & !is.na(expected) ~ "TP",
    type == "Mock" & is.na(Abundance) & expected > 0 ~ "FN"
  )) 

# Plot tree
keeptips <- heatmap_dat %>% 
  filter(Abundance > 0) %>%
  dplyr::select(OTU, Species) %>%
  distinct()

tree <- phy_tree(ps_run1)
tree <- drop.tip(tree, tree$tip.label[!tree$tip.label %in% keeptips$OTU])

tree$tip.label <- tree$tip.label %>%
  tibble::enframe(name=NULL, value="OTU") %>%
  left_join(keeptips) %>%
  pull(Species)

mock_tree <- ggtree(tree, ladderize = TRUE)
library(ggnewscale)

gg.run1_heatmap <- heatmap_dat %>%
  left_join(mock_tree$data %>% dplyr::rename(Species = label)) %>%
  mutate(Species = factor(Species)) %>%
    ggplot(aes(x=sample_name, y=fct_reorder(Species, y), fill=Abundance)) +
    geom_tile() +
  scale_fill_viridis_c(labels = scales::percent, na.value = NA, alpha=0.9) +
  new_scale_fill() +
  geom_tile(aes(fill = outcome))+
  scale_fill_manual(values=c("FP" = "#e41a1c", "TP" = NA, "FN"= "grey60")) +
  base_theme+
  theme(legend.position = "right",
        axis.text.y = element_text(face="italic"),
        axis.title.y = element_blank())+
    labs(x="Sample",
         y="Taxon") +
  facet_grid(~pcr_primers,scales="free_x", drop=TRUE)

mock_tree + gg.run1_heatmap + plot_layout(widths = c(1,4))
```

## Bias

```{r Drosophila bias}
ps_bias <- ps_run1

exp <- read_csv("sample_data/expected_quant_merged.csv") %>%
  dplyr::rename(sample_name = Sample) %>%
  pivot_longer(-sample_name,
               names_to= "taxon",
               values_to= "expected") %>%
  mutate(taxon = str_replace(taxon, pattern=" ",replacement="_"),
         sample_name = str_remove(sample_name, "-exp*.$")) %>%
  dplyr::filter(!is.na(sample_name),
  !str_detect(sample_name, "CM[0-9]|CT[0-9]|CML[0-9]|Syn"),
  expected > 0) %>%
  distinct()

#plot expected
exp %>% 
  dplyr::filter(str_detect(sample_name, "^D")) %>%
  group_by(sample_name) %>%
  mutate_at(vars(expected), ~ . / sum(.) ) %>% #Convert to proportions
  ggplot(aes(x=sample_name, y=expected, fill=taxon)) +
  geom_col(position="stack") + 
  scale_fill_brewer(palette="Spectral") +
  scale_y_continuous(labels=scales::percent) +
  base_theme+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(x = "Sample Name",
       y= "Relative abundance",
       title="Expected mock communities",
       fill="Species")

#Get observed
sam <- speedyseq::psmelt(ps_bias) %>%
  janitor::clean_names() %>%
  mutate(taxon = species) %>%
  filter(!str_detect(genus, "__")) %>%#Remove unclassified
  dplyr::select(sample_name, taxon, abundance, pcr_primers, fcid, material_type = type) %>%
  mutate(sample_name = sample_name %>%
           str_remove_all("BF1-BR1-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-|fwhF2-fwhR2n-") %>%
           str_remove("HLVKYDMXX_"),
         sample_id = paste0(fcid, "_", sample_name),
         sample_name = sample_name %>%
           str_remove_all("HLVKYDMXX_|CK3HD_|CB3DR_|CJKFJ_")) 

#Join tables 
joint <- sam %>%
  filter(taxon %in% exp$taxon,
         sample_name %in% exp$sample_name) %>%
  left_join(exp, by = c("sample_name","taxon")) %>%
  dplyr::rename(observed = abundance) %>%
  filter(observed > 0) %>%
  mutate(expected = replace_na(expected, 0),
         observed0 = (observed + 0.5) * (expected > 0),
         error = observed0 / expected) %>%
  distinct() %>%
  group_by(sample_id) %>%
  mutate_at(vars(observed, expected), ~ . / sum(.) ) %>% #Convert to proportions
  group_by(sample_id) %>%
  group_modify(~{
    clr_denom <- .x %>%
      pull(error) %>%
      gm_mean(na.rm = TRUE)
    .x %>%
      mutate(error_clr = log(error / clr_denom))
  })


# Visualise the error in all pairwise ratios
gg.ratio <- joint %>%
  dplyr::filter(expected > 0 ) %>%
  dplyr::mutate(Taxon = taxon %>%
                 str_replace("^.+_", paste0(str_extract(taxon, "."), "_"))) %>%
    compute_ratios(group_vars = c("sample_id", "material_type", "pcr_primers")) %>%
  mutate(Pair = paste(Taxon.x, Taxon.y, sep = ":")) %>% 
  ggplot(aes(Pair, expected, colour=sample_id)) +
  geom_hline(yintercept = 1, alpha=0.8) +
  geom_jitter(alpha=0.7) +
  scale_y_log10() +
  base_theme+
  facet_grid(pcr_primers~.)+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=45, hjust=1))+
  labs(y= "Error in taxon ratios (log10)") 

gg.ratio

# Estimate bias on geometrically centred ratios
bias <- joint %>%
  filter(material_type=="DrosMock") %>%
  filter(expected > 0) %>%
  group_by(pcr_primers, material_type)  %>%
    nest() %>%
    mutate(fit = purrr::map(data, ~lm(error_clr ~ 0 + taxon , data = .)), 
           tidied =  purrr::map(fit, broom::tidy),
           aug =  purrr::map(fit, broom::augment)
    ) %>%
  dplyr::select(-fit) %>%
  unnest(tidied) %>%
  mutate(taxon = str_remove(term, "taxon"),
         estimate = exp(estimate)) %>%
   dplyr::select(pcr_primers, material_type, taxon, estimate)


# Bootstrap estimation
set.seed(606)
boot_models <- joint %>%
  filter(material_type=="DrosMock") %>%
  filter(expected > 0) %>%
  group_by(pcr_primers) %>%
  group_modify(~{
    bootstraps(.x, times=1000, apparant=TRUE)  %>%
    mutate(fit =  purrr::map(splits, ~lm(error_clr ~ 0 + taxon , data = .)), 
           coef =  purrr::map(fit, function(x){
             broom::tidy(x) %>% 
               mutate(estimate = exp(estimate))})) %>%
    int_pctl(coef) 
    }) %>%
  mutate(taxon = str_remove(term, "taxon")) %>%
  dplyr::select(pcr_primers, taxon, lower = .lower, estimate = .estimate, upper = .upper)

# See how well bias esitmate fits data
preds <- joint %>%
  left_join(boot_models) %>%
  dplyr::mutate(predicted = expected * estimate) %>%
  group_by(sample_id) %>%
  mutate_at(vars(predicted), ~ . / sum(., na.rm=TRUE) ) %>%
  filter(expected > 0) %>%
  filter(observed > 0)

# Get RMSE
errors <- preds %>%
  group_by(pcr_primers) %>%
  rmse(truth = observed, estimate = predicted) %>%
  dplyr::select(pcr_primers, rmse = .estimate)

gg.bias_fits <- preds %>%
  mutate(taxon = str_replace(taxon, "_", " ")) %>%
  left_join(errors) %>%
  mutate(rmse = paste0("RMSE = ", round(rmse, 2))) %>%
  ggplot(aes(x=logit(predicted), y=logit(observed), color = taxon))+
  geom_abline(intercept = 0, slope = 1, color = "grey") +
    geom_jitter(width = 0.1, height = 0, alpha=0.7, size=2) +
  geom_text(aes(x=-6, y=-1, label=rmse), check_overlap = TRUE, inherit.aes = FALSE)+
  facet_wrap(~pcr_primers)+
  coord_fixed()+
  scale_color_brewer(palette = "Paired")+
  base_theme +
  theme(
        panel.spacing.x = unit(1, "lines"),
        legend.position = "bottom",
        legend.text = element_text(face = "italic", size=10)
    ) +
   labs(x = "log-odds(Predicted proportion)", 
     y = "log-odds(Observed proportion)",
     colour = "Taxon") 


gg.bias_fits

# Write out bias explained for supplementary 
pdf(file="fig/supplementary/bias_model_fit.pdf", width = 11, height = 8 , paper="a4r")
  plot(gg.bias_fits)
try(dev.off(), silent=TRUE)


# Plot bias estimates
gg.bias <- preds %>%
  ungroup() %>%
  dplyr::select(taxon, estimate, upper, lower, pcr_primers) %>%
  distinct()%>%
  left_join(mock_tree$data %>% dplyr::mutate(taxon = label %>% str_replace(" ", "_"))) %>%
  mutate(taxon = taxon %>% 
           str_replace("^Drosophila_", "D. ")%>%
           str_replace("^Scaptodrosophila_", "Sca. ")) %>%
  mutate(taxon = factor(taxon))%>%
  ggplot(aes(x = fct_reorder(taxon, y), y = estimate, colour = pcr_primers, group = pcr_primers)) +
  geom_hline(yintercept = 1, alpha=0.5, colour = "grey20") +
  geom_pointrange(aes(ymin = lower, ymax = upper), position=position_dodge(width=0.6))+
  #coord_flip() +
  scale_y_log10()+
  #facet_grid(pcr_primers~.) +
  base_theme+
  scale_colour_brewer(palette="Paired") +
  labs(
    x = NULL,
    colour="PCR primers",
    y = "Efficiency / Geometric mean") +
  theme(axis.text.x = element_text(face = "italic"),
        legend.position = "bottom")

gg.bias

#bias is missing for BF1 biarmipes because it was low reads
```

## Multifig1

Comparison of primers for bias, off-target amp, and PCA of overlap

```{r run 1 multifig}

fig1a <- mock_tree + gg.run1_heatmap + plot_layout(widths = c(1,6))

fig1 <- fig1a / gg.bias + plot_layout(heights = c(2,1))+ plot_annotation(tag_levels = 'A') & theme(plot.margin = unit(c(1,1,1,1), "mm"))

fig1
# Write out figure 1
pdf(file="fig/run1_multifig.pdf", width = 11, height = 8 , paper="a4r")
  plot(fig1)
try(dev.off(), silent=TRUE)
```



## Classification summary

The ratio of root assigned reads to those under species level can be used as a proxy for taxonomic assignment

```{r, run 1 tax summary}
# Subset to run 1
ps <- readRDS("output/rds/ps_idtaxaExact.rds")%>%
  subset_samples(fcid %in% c("CB3DR")) 

tax_summary <- speedyseq::psmelt(ps) %>%
  dplyr::select(pcr_primers, sample_id, rank_names(ps), Abundance) %>%
  pivot_longer(rank_names(ps),
               names_to="rank",
               values_to="name") %>%

  mutate(name = replace(name, str_detect(name, "__"), NA),
         rank = str_replace(rank, "Kingdom", "Root")) %>% 
    mutate(detection = case_when(
    Abundance > 0 ~ 1,
    Abundance == 0 ~ 0
  )) %>%
  group_by(pcr_primers, rank) %>% 
  #dplyr::filter(!is.na(name)) %>%
  dplyr::summarise(reads_classified = sum(Abundance * !is.na(name)),
                   asvs_classified = sum(detection * !is.na(name)))

# Lineplot
tax_plot_data <- tax_summary %>%
  ungroup %>%
  filter(!rank=="Root") %>%
  left_join(tax_summary %>% 
              filter(rank=="Root") %>%
              dplyr::select(pcr_primers, root_reads = reads_classified, root_asv = asvs_classified)) %>%
  mutate(read_prop = reads_classified / root_reads,
         asv_prop = asvs_classified / root_asv)%>% 
  dplyr::select(pcr_primers, rank, contains("prop"))  %>%
  pivot_longer(contains("prop"),
               names_to="type",
               values_to = "value") %>%
  mutate(type = type %>% 
           str_replace("asv_prop", "ASVs classified") %>%
           str_replace("read_prop", "Sequences classified")) %>%
  mutate(rank = factor(rank, levels = rank_names(ps)[2:length(rank_names(ps))]))

gg.reads_classified <- tax_plot_data  %>%
  ggplot(aes(x=rank, y=value, fill = pcr_primers, group=pcr_primers)) +
  geom_col(width=0.7, position=position_dodge(width = 0.8)) +
  #geom_text(aes(label = percent(value, accuracy = 0.1)),
  #          position=position_dodge(width = 0.8), angle=270,hjust =-0.1, colour="white", size=4 )+
  base_theme +
  facet_wrap(~type, nrow = 2)+
  #facet_grid(~type, cols=1, rows=2) +
  scale_fill_brewer(palette="Paired") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL,
       y=NULL)


gg.reads_classified
```


# Supplementary Comparison of primer replicates for detection and bias

# Run 2 - comparison between replicate primers

Run 2 - see the differences in bias between replicates

```{R run2}
# RUN 2
ps_run2 <- readRDS("output/rds/ps_idtaxaExact.rds") %>%
  subset_samples(fcid %in% c("CK3HD")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE) %>%
  speedyseq::tax_glom(taxrank="Species")

exp <- read_csv("sample_data/expected_quant_merged.csv") %>%
  dplyr::rename(sample_name = Sample) %>%
  pivot_longer(-sample_name,
               names_to= "taxon",
               values_to= "expected") %>%
  mutate(taxon = str_replace(taxon, pattern="_",replacement=" "),
         sample_name = str_remove(sample_name, "-exp*.$")) %>%
  dplyr::filter(!is.na(sample_name),
  !str_detect(sample_name, "CM[0-9]|CT[0-9]|CML[0-9]|Syn"),
  expected > 0) %>%
  distinct()

#Get observed

sam <- ps_run2%>%
  speedyseq::psmelt() %>%
  janitor::clean_names() %>%
  mutate(taxon = species) %>%
  filter(!str_detect(genus, "__")) %>%#Remove unclassified
  dplyr::select(sample_name, sample_id, amp_rep, taxon, abundance, pcr_primers, fcid, material_type = type) %>%
  mutate(sample_name = sample_name %>%
           str_remove_all("BF1-BR1-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-|fwhF2-fwhR2n-") %>%
           str_remove_all("HLVKYDMXX_|CK3HD_|CB3DR_|CJKFJ_"))

#Join tables 
joint <- sam %>%
  filter(taxon %in% exp$taxon,
         sample_name %in% exp$sample_name,
         abundance > 0) %>%
  left_join(exp, by = c("sample_name","taxon")) %>%
  dplyr::rename(observed = abundance) %>%
  mutate(expected = replace_na(expected, 0),
         observed0 = (observed + 0.5) * (expected > 0),
         error = observed0 / expected) %>%
  distinct() %>%
  group_by(sample_id) %>%
  mutate_at(vars(observed, expected), ~ . / sum(.) ) %>% #Convert to proportions
  ungroup() %>%
  group_by(sample_id) %>%
  group_modify(~{
    clr_denom <- .x %>%
      pull(error) %>%
      gm_mean(na.rm = TRUE)
    .x %>%
      mutate(error_clr = log(error / clr_denom))
  })

# Bootstrap estimation
set.seed(606)
boot_models <- joint %>%
  filter(material_type=="DrosMock") %>%
  filter(expected > 0) %>%
  group_by(pcr_primers, amp_rep) %>%
  group_modify(~{
    bootstraps(.x, times=1000, apparant=TRUE)  %>%
    mutate(fit =  purrr::map(splits, ~lm(error_clr ~ 0 + taxon , data = .)), 
           coef =  purrr::map(fit, function(x){
             broom::tidy(x) %>% 
               mutate(estimate = exp(estimate))})) %>%
    int_pctl(coef) %>%
    mutate(taxon = str_remove(term, "taxon"))
    }) %>%
  dplyr::select(pcr_primers, amp_rep, taxon, lower = .lower, estimate = .estimate, upper = .upper)

# See how well bias estiamte fits data
preds <- joint %>%
  left_join(boot_models) %>%
  dplyr::mutate(predicted = expected * estimate) %>%
  group_by(sample_id) %>%
  mutate_at(vars(predicted), ~ . / sum(., na.rm=TRUE) ) %>%
  filter(expected > 0) %>%
  filter(observed > 0)

# Plot bias estimates for the 3 separate primers
gg.biasreps <- preds %>%
  ungroup() %>%
  dplyr::select(taxon, estimate, upper, lower, pcr_primers, amp_rep) %>%
  distinct()%>%
  mutate(amp_rep = factor(amp_rep, levels = c("1","2","3"))) %>%
  mutate(taxon = str_replace(taxon, "^.*_", "D. ")) %>%
  ggplot(aes(x = taxon, y = estimate, colour = amp_rep, group =amp_rep)) +
  geom_hline(yintercept = 1, alpha=0.5, colour = "grey20") +
  geom_pointrange(aes(ymin = lower, ymax = upper), position=position_dodge(width=0.5))+
  facet_grid(~pcr_primers) +
  coord_flip() +
  base_theme+
  scale_y_log10()+
  scale_colour_brewer(palette="Paired") +
  labs(
    x = NULL,
    y = "Efficiency / Geometric mean",
    colour = "Tagged PCR primer") +
  theme(axis.text.y = element_text(face = "italic"),
        legend.position = "bottom")

gg.biasreps

# Write supplementary figure
pdf(file="fig/supplementary/run2_biasreps.pdf", width = 11, height = 8 , paper="a4r")
  plot(gg.biasreps)
try(dev.off(), silent=TRUE)
```

# Trap catch analysis

## Insects trapped summary

```{r Trapped insects}
trapdat <- read_csv("Fieldwork/trap_data_clean.csv") %>%
  janitor::clean_names()%>%
  mutate(catch = as.numeric(catch)) %>%
  mutate(orchard_type = case_when(
         str_detect(fruit, "Peach") ~ "Stonefruit",
         str_detect(fruit, "Nectarine") ~ "Stonefruit",
         str_detect(fruit, "Cherries") ~ "Cherries",
         TRUE ~ "Cherries"
  )) %>%
  dplyr::filter(!attractant=="Sachet") %>%
  dplyr::select(-treatment)


# Average catch
trapdat %>%
  group_by(week, attractant, orchard, orchard_type) %>%
  dplyr::summarise(catch = sum(catch, na.rm=TRUE))
  
##Total catch by orchard and week
gg.trapweek <- trapdat %>% 
  mutate(week = as.factor(week)) %>%
  ggplot(aes(x=week, y=catch, colour=attractant, group=week))+
  geom_jitter(size=2, alpha=0.6, width=0.2, height=0.2)+
  geom_boxplot(fill=NA, outlier.colour = NA)+
  facet_grid(attractant~orchard_type, scales="free") +
  base_theme +
  scale_colour_brewer(palette="Paired") 

##Total catch by orchard and trap type
gg.trapall <- trapdat %>% 
  ggplot(aes(x=attractant, y=catch, colour=attractant))+
  geom_jitter(size=2, alpha=0.6, width=0.2, height=0.2)+
  geom_boxplot(fill=NA, outlier.colour = NA)+
  facet_grid(~orchard_type) +
  base_theme +
  scale_colour_brewer(palette="Paired")  + 
  labs(x = "Collection method",
       y = "Estimated trap catch")


```

## Subset to just run 3

```{r subset}
ps_run3 <- ps2 %>%
  subset_samples(fcid %in% c("HLVKYDMXX")) %>%
  subset_taxa(Phylum =="Arthropoda") %>%
  #prune_taxa(!str_detect(Species, "Synthetic")) %>%
  filter_taxa( function(x) mean(x) > 0, TRUE)

ps_run3 <-  prune_samples(sample_sums(ps_run3) >0 , ps_run3)

# Need to remove synthetics
```

## Alpha diversity metrics

No significant differences

```{r alpha div}
dir.create("output/alpha")

# Get a histogram of taxon sums
taxa_sums(ps_run3) %>%
  as_tibble(rownames = "OTU") %>%
  filter(value > 0) %>%
  ggplot(aes(x = value)) +
  geom_histogram() 

# Get richness measures
richness <- phyloseq::estimate_richness(ps_run3, measures=c("Chao1", "Shannon", "Simpson")) %>% #
  rownames_to_column("sample_name") %>%
  mutate(sample_name = str_replace(sample_name, "\\.", "-"))

#Richness is giving a warning because its acting on the trimemd dataset

#Set number of randomisations for calculating significance
# Calculate Faith's PD-index & Species richness - with Standard errors
sespd <- picante::ses.pd(as(phyloseq::otu_table(ps_run3), "matrix"),  phyloseq::phy_tree(ps_run3), null.model = "taxa.labels", include.root = F, runs = 9)

# Join together
div_table <- sespd %>%
  rownames_to_column("sample_name") %>%
  dplyr::select(sample_name, ntaxa, pd.obs) %>%
  dplyr::rename(pd = pd.obs, alpha = ntaxa) %>%
  left_join(richness, by="sample_name") %>%
  left_join(sample_data(ps_run3) %>% 
              as("data.frame") %>%
              filter(!duplicated(sample_id)) %>%
              dplyr::select(amp_rep, sample_name = sample_id, type),
            by = "sample_name") 

# Filter to just the field collected types
div_table <- div_table %>%
  filter(type %in% c("DC", "ACV", "FF", "SPD"))

# Summarise means
div_table %>%
  summarise(across(where(is.numeric), mean, na.rm=TRUE))

# Difference in alpha diversity between replicates
#report::report(aov(alpha ~amp_rep, data=div_table))
#report::report(aov(Shannon ~amp_rep, data=div_table))
#report::report(aov(pd ~amp_rep, data=div_table))
#
# Difference in alpha diversity between trap types
report::report(aov(alpha ~type, data=div_table))
broom::tidy(TukeyHSD(aov(alpha ~type, data=div_table)))

report::report(aov(Shannon ~type, data=div_table))
broom::tidy(TukeyHSD(aov(Shannon ~type, data=div_table)))

report::report(aov(pd ~type, data=div_table))
broom::tidy(TukeyHSD(aov(pd ~type, data=div_table)))

# SHould do on the original data without merging by spp

# plot differences in alpha diversity
gg.alpha <- div_table %>% 
  pivot_longer(2:7,
               names_to="metric",
               values_to="value") %>%
  filter(metric %in% c("alpha", "pd", "Shannon")) %>%
  mutate(metric = metric %>% 
           str_replace("alpha", "ASV Richness") %>%
           str_replace("pd", "Phylogenetic Diversity") %>%
           str_replace("Shannon", "Shannon Index")) %>%
  ggplot(aes(x=type, y=value, colour=type)) +
  geom_jitter(size=2, alpha=0.6, width=0.2, height=0)+
  geom_boxplot(fill=NA)+
  coord_flip()+
  base_theme+
  scale_colour_brewer(palette="Paired") +
  facet_wrap(.~metric,ncol=1, scales="free")

gg.alpha

pdf(file="fig/supplementary/trap_catches.pdf", width = 8, height = 6 , paper="a4r")
  plot(gg.trapweek)
  plot(gg.trapall)
  plot(gg.alpha)
try(dev.off(), silent=TRUE)

# This could be nice really small, under the total catch

```

## Beta diversity metrics

```{r Distances}
# Get OTU tables
otutab <- otu_table(ps_run3) %>%
  as("matrix")
#Impute zeroes for compositional distances
otutab_n0 <- as.matrix(zCompositions::cmultRepl(otutab, method="CZM", output="p-counts"))
#Root phylogenetic tree
phy_tree(ps_run3) <- multi2di(phy_tree(ps_run3))
phy_tree(ps_run3) <- makeNodeLabel(phy_tree(ps_run3), method="number", prefix='n')
name.balance(phy_tree(ps_run3), tax_table(ps_run3), 'n1') #Get root

#Calculate different distance metrics
metrics <- c("Bray", "Jaccard", "Aitchison", "Philr", "Unifrac", "WUnifrac")
distlist <- vector("list", length=length(metrics))
names(distlist) <- metrics

distlist$Jaccard <- as.matrix(vegdist(otutab, method="jac",binary = T))
distlist$Bray <- as.matrix(vegdist(otutab, method="bray"))
distlist$Aitchison <- as.matrix(vegdist(CoDaSeq::codaSeq.clr(otutab_n0), method="euclidean"))
distlist$Philr <- as.matrix(vegdist(philr::philr(otutab_n0, phy_tree(ps_run3),
                                                part.weights='enorm.x.gm.counts',
                                                ilr.weights='blw.sqrt'), method="euclidean"))
distlist$Unifrac <- as.matrix(phyloseq::UniFrac(ps_run3, weighted=FALSE, parallel = TRUE))
distlist$WUnifrac <- as.matrix(phyloseq::UniFrac(ps_run3, weighted=TRUE, parallel = TRUE))

# Probably only need to use philr dist
```

# Adonis and betadisper

```{r betatest}
# Adonis test
metadata <- sample_data(ps_run3) %>%
  as("data.frame")

# Test difference by community type
adonis_results <- distlist %>%
  purrr::map(function(x) {
    bind_rows(
    broom::tidy(adonis(x~type, method="euclidean", data=metadata)$aov.tab) #%>% dplyr::slice(1)
    )
})  %>%
  bind_rows(.id="dist")

# Check homogeneity
betadisper_results <- distlist %>%
  purrr::map(function(x) {
    y <- as.dist(x[metadata$sample_id, metadata$sample_id])
  bind_rows(
    as.data.frame(permutest(vegan::betadisper(y, metadata$type))$tab) %>%
      #dplyr::slice(1) %>% 
      mutate(term="type")
  )
})  %>%
  bind_rows(.id="dist")

dir.create("output/beta")
write_csv(adonis_results, "output/beta/adonis.csv")
write_csv(betadisper_results, "output/beta/adonis.csv")
```

# PCA plots

```{r pca plots}
# Get philr distance
philr.dist <- as.dist(distlist$Philr)

#Philr PCOA
philr.pcoa <- ordinate(ps_run3, 'PCoA', distance=philr.dist)

gg.pca <- plot_ordination(ps_run3, philr.pcoa, color = "type") +
  geom_point(size=3, alpha=0.8) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.5) +  
  geom_vline(xintercept = 0, linetype=2, alpha=0.5) +
  base_theme +
  #coord_fixed() +
  scale_colour_brewer(palette="Paired") +
  theme(legend.position = "bottom") +
  labs(colour = "Sample Type")

gg.pca

# Heirarchial clustering
dend <- hclust(philr.dist, method="average")

p3 <- ggtree(as.phylo(dend) ) + 
  theme_tree2()

colours_p3 <- p3$data %>%
  left_join(as_data_frame(sample_data(ps_run3)) %>%
              dplyr::select(sample_id, type) %>%
  dplyr::rename(label = sample_id)
    )

p3 <- p3 %<+% colours_p3  + 
  geom_tippoint(aes(colour=as.factor(type)))  + 
  geom_tiplab(aes(colour=as.factor(type)))+
  scale_colour_brewer(palette="Paired") +
    scale_x_continuous(expand=c(0, 30)) +
  theme(legend.position = "none")

# Plot together
p3 + gg.pca


```

# Plot a heatmap of insect distances vs clustering of samples

```{r Heatmap}
#Prepare co-occurance matrix
coocur <- ps_run3 %>%
    otu_table %>%
    as("matrix")

# Sample tree 
h_tree <- as.phylo(dend)

# P cophenetic distance
s_tree <- phy_tree(ps_run3)

s_tree <- drop.tip(s_tree, setdiff(s_tree$tip.label, colnames(coocur)))
coocur <- coocur[h_tree$tip.label, s_tree$tip.label]

#Could do the imputing on coocur

# Cophyloplot
coocur.lut <- which(coocur >0, arr.ind=TRUE)
assoc <- cbind(rownames(coocur)[coocur.lut[,1]], colnames(coocur)[coocur.lut[,2]])
# Rotate the nodes using phytools
obj <- cophylo(tr1=h_tree, tr2=s_tree, assoc=assoc, rotate=TRUE) 

# Extract the goods for ggtree
tree1 <- obj[["trees"]][[1]]
p1 <- ggtree(tree1, ladderize=FALSE) + geom_tiplab()

# OTU tree
tree2 <- obj[["trees"]][[2]]
p2 <- ggtree(tree2, ladderize=FALSE) + geom_tiplab()

# Tanglegram
dd <- obj$assoc %>%
  as_data_frame() %>%
  magrittr::set_colnames(c("label.x", "label.y")) %>%
  left_join(p1$data %>% dplyr::select(label, y) %>% dplyr::rename(label.x = label), by="label.x") %>%
  left_join(p2$data %>% dplyr::select(label, y) %>% dplyr::rename(label.y = label), by="label.y") %>%
  rownames_to_column("assoc")

gg.heatmap <- dd %>%
  dplyr::rename(Sample_Name = label.x,
                OTU = label.y,
                pos_x = y.x,
                pos_y = y.y) %>%
  dplyr::select(Sample_Name, OTU, pos_x, pos_y) %>%
  dplyr::mutate(Sample_Name = factor(Sample_Name),
                Sample_Name = fct_reorder(Sample_Name, pos_x),
                OTU = factor(OTU),
                OTU = fct_reorder(OTU, pos_y)) %>%
   ggplot(aes(x=Sample_Name, y=OTU)) +
  geom_tile() +
  base_theme+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") 

#Adjusted trees - Sca

top <- wrap_elements(grid::textGrob('')) +(p1+ coord_flip() + scale_x_reverse(expand=c(0,0))+ scale_y_continuous(expand=c(0,0))) + plot_layout(widths=c(1,3)) 
bottom <- p2+ scale_y_continuous(expand=c(0,0)) +gg.heatmap + plot_layout(widths=c(1,3))

gg.treemap <- top / bottom + plot_layout(heights=c(1,3))                                                         

gg.treemap

gg.heatmap <- psmelt(ps_run3) %>%
  dplyr::select(Abundance, label = sample_id, OTU, rank_names(ps_run3)) %>%
  left_join(p3$data) %>%
  group_by(label) %>%
  mutate(Abundance = replace(Abundance, Abundance==0, 0.5)) %>%
  mutate(Abundance = metacal::clr(Abundance, na.rm = TRUE)) %>%
  ungroup %>%
  ggplot(aes(x = label, y = Species, fill = Abundance )) +
  geom_tile() +
  base_theme
  
(p3 + coord_flip() + scale_x_reverse()) / gg.heatmap

```

# Phylogeny of trap catches

```{r Phylogeny}

taxa_names(ps_run3) <- taxa_names(ps_run3)%>% str_remove("^.*-")

tree <- phy_tree(ps_run3)

# Colour tree by order
#


p1dat <- speedyseq::psmelt(ps_run3) %>%
  dplyr::select(sample_id, OTU, type, Abundance) %>%
  #group_by(sample_id) %>%  
  #mutate_at(vars(Abundance), ~ . / sum(.)) %>%
  filter(Abundance > 0) %>%
  drop_na() %>%
  dplyr::filter(type %in% c("ACV", "DC", "FF", "SPD")) %>%
  group_by(OTU, type) %>%
  summarise(Abundance = sum(Abundance)) %>%
  distinct() %>%
  pivot_wider(names_from = type,
              values_from = Abundance,
              values_fill = 0) %>%
  column_to_rownames("OTU")

tree <- drop.tip(tree, tree$tip.label[!tree$tip.label %in% rownames(p1dat)])

circ <- ggtree(tree, layout = "circular", branch.length = 'none')

p1 <- gheatmap(circ, p1dat, offset=0, width=.4, color="gray20",
               colnames_angle=95, colnames_offset_y = 0) +
    scale_fill_viridis_c(option="D", name="Abundance", alpha=0.9, trans="log10") 

# Also try add week? and orchard?
library(ggnewscale)
p2dat <- speedyseq::psmelt(ps_run3) %>%
  mutate(orchard = case_when(
    str_detect(sample_name,"^T") ~ "Stonefruit",
    str_detect(sample_name,"^M") ~ "Cherries",
    TRUE ~ as.character(NA)
  )) %>%
  dplyr::select(OTU, orchard, Abundance) %>%
    filter(Abundance > 0) %>%
  drop_na() %>%
  group_by(OTU, orchard) %>%
  summarise(Abundance = sum(Abundance)) %>%
  distinct() %>%
  pivot_wider(names_from = orchard,
              values_from = Abundance,
              values_fill = 0) %>%
  column_to_rownames("OTU")
  
p2 <- p1 + new_scale_fill()
gg.phylo <- gheatmap(p2, p2dat, offset=5.5, width=.2,
         colnames_angle=90, colnames_offset_y = .25, color="gray20") +
    scale_fill_viridis_c(option="B", name="Abundance", trans="log10")+
  geom_tiplab(offset=9, align=FALSE, size=3) 

#p3 <- p2 + new_scale_fill()
#gg.phylo <- gheatmap(p3, p2dat %>% dplyr::select(Stonefruit), offset=7, width=.2,
#         colnames_angle=90, colnames_offset_y = .25, color="gray20") +
#    #scale_fill_gradient(low="sandybrown", high="firebrick", trans="log10")+
#    scale_fill_viridis_c(option="B", name="Abundance", trans="log10")+ 
#    geom_tiplab(offset=12, align=FALSE, size=3) 
#
gg.phylo

# Write out bias explained for supplementary 
pdf(file="fig/phylogeny_occurance.pdf", width = 11, height = 8 , paper="a4r")
  plot(gg.phylo)
try(dev.off(), silent=TRUE)
  
# Colour tree by Order - should be able to do this with highlight clade
# need to make the tree better though

# Work out how to properly compositionall merge a sample

```

## Detection probability

How do the primers affect its probability of detection (presence/absence). We used a generalized linear mixed effects model of presence/absense as a function of primer primer. To account for the paired design (where both types of samples were derived from the same community), we defined Sample ID as a random effect. Since the response was binomial, we assumed a logit‐link and binomially distributed error

```{r detection probability}
# Use joint to get a table of expected/absent
det_table <- joint %>% 
  filter(material_type=="DrosMock") %>%
  mutate(detected = case_when(
    expected > 0 & observed > 0 ~ 1,
    expected > 0 & observed == 0 ~ 0,
    TRUE ~ as.numeric(NA)
    )) %>%
  filter(!is.na(detected)) %>% #Still some false positives remeaining!
  mutate(extract_id = sample_name %>%
           str_remove_all("BF1-BR1-|fwhF2-fwhR2n-|SauronS878-HexCOIR4-|fwhF2-HexCOIR4-"))

# Estimate detection efficiency of all taxa - Better done within tidymodels?
library(lme4)
detmodel <- det_table %>%
  filter(fcid=="CB3DR") %>%
  dplyr::select(sample_id, extract_id, taxon, detected, pcr_primers) %>%
  #group_by(taxon) %>%
  nest() %>%
  mutate(fit = map(data, ~glmer(detected ~ pcr_primers + (1|extract_id), family=binomial, data = .)), 
           tidied = map(fit, broom::tidy),
           aug = map(fit, broom::augment)

  ) %>%
  dplyr::select(-fit) %>%
  unnest(tidied) 

# Estimate detection efficiency of targets - Doesnt work because they were pretty much all detected!
library(lme4)
det_suzukii <- det_table %>%
  filter(fcid=="CB3DR") %>%
  group_by(taxon) %>%
  nest() %>%
  mutate(fit = map(data, ~glmer(detected ~ pcr_primers + (1|extract_id), family=binomial, data = .)), 
           tidied = map(fit, broom::tidy),
           aug = map(fit, broom::augment)

  ) %>%
  dplyr::select(-fit) %>%
  unnest(tidied) 

det_suzukii
```

# Reproducability Receipt

```{details, echo = FALSE, details.summary = 'Reproducability receipt'}
# datetime
Sys.time()
#repository
git2r::repository()
sessioninfo::session_info()
```
